---
title: "SCALE Study"

subtitle: "Tidy-up the shapefiles for 72 clusters"

author: |
  | Peter MacPherson
  |
  | Liverpool School of Tropical Medicine, Liverpool, UK
  | Malawi-Liverpool-Wellcome Clinical Research Programme, Blantyre, Malawi
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    theme: cosmo
    highlight: espresso
    toc: true
---

<br>

## 1. Backgound

Use this script to clean the SCALE Study clusters for future studies

<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, echo=FALSE, include=FALSE, message=FALSE, comment=NA, warning=FALSE}
library(tidyverse)    #for data manipulation
library(sf)
library(lwgeom)
library(ggforce)
library(spdep)
library(rgeos)
library(knitr)
library(here)

```

<br>

## 3. Import datasets

Import data required for the analysis.

```{r import}

raw_scale_clusters <- st_read(here::here("data-raw/01_scale_clusters/72_clusters_polyg_scale.shp"))

#add an ID for each cluster
raw_scale_clusters <- raw_scale_clusters %>%
  mutate(cluster = paste0("c",row_number())) %>%
  dplyr::select(cluster, geometry)


ggplot(raw_scale_clusters) +
  geom_sf() +
  theme_minimal()

```

<br>

## 4. Tidying and recoding

try tidying up the slivers

```{r tidy}

#what is the crs for object currently
st_crs(raw_scale_clusters)

#what attributes does it have?
print(head(raw_scale_clusters))

#Is a single polygon, so needs to be converted to a MULTIPOLYGON
raw_scale_clusters <- st_cast(raw_scale_clusters, "MULTIPOLYGON")

#did that work?
print(head(raw_scale_clusters))
st_is_valid(raw_scale_clusters, reason = TRUE)

#So there are some non-valid clusters, because boundaries self-intersect, or have slivers
#Try to make these polygons valid
raw_scale_clusters <- st_make_valid(raw_scale_clusters)
st_is_valid(raw_scale_clusters)

#Looks like that fixed it.

ggplot(raw_scale_clusters) +
  geom_sf() +
  theme_minimal()

#write to shapefile for checking slivers
#st_write(raw_scale_clusters, "raw_scale_cluster.shp")

#There are still some slivers
raw_scale_clusters <- raw_scale_clusters %>%
  st_buffer(.00001) 

#write to shapefile for checking slivers
#st_write(raw_scale_clusters, "raw_scale_cluster2.shp")

#did that work?
print(head(raw_scale_clusters))
st_is_valid(raw_scale_clusters, reason = TRUE)

#covert back into a MULTIPOLYGON
raw_scale_clusters <- st_cast(raw_scale_clusters, "MULTIPOLYGON")

#write to shapefile for checking slivers
#st_write(raw_scale_clusters, "raw_scale_cluster3.shp")

#By checking on QGIS - boundaries are now resolved

scale_72_clusters <- raw_scale_clusters

```

Plot with labels


```{r}

ggplot(scale_72_clusters) +
  geom_sf() +
  geom_sf_text(aes(label=cluster), colour="blue") +
  theme_minimal() +
  labs(title="SCALE Study: 72 clusters",
       x="",
       y="")

ggsave(here::here("plots/scale_72_clusters.pdf"), dpi=600)

use_data(scale_72_clusters, overwrite = TRUE)

```





Now check by looking for adjacency of boundaries

```{r}


sy2 <- poly2nb(as(scale_72_clusters, "Spatial"), row.names = scale_72_clusters$cluster, queen=FALSE, snap = 0.0001)
summary(sy2)

head(sy2)

plot(st_geometry(scale_72_clusters), borders="grey")
plot(sy2, coordinates(as(scale_72_clusters, "Spatial")), col="darkgreen", add=TRUE)
```


<br>

## 5. Reproducibility

This reproduction of the analysis was run by: 

```{r sysinfo, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

sysinfo <- Sys.info()

sysinfo <- data.frame(keyName=names(sysinfo), value=sysinfo, row.names=NULL)

sysinfo %>% kable()
```

Analysis was run at **`r Sys.time()`**, and using the following Session Info:

```{r sessioninfo, echo=FALSE, results='markdown', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```
